Incremental tactics

### 2012-05-07 doc/papers/LFMTP2012/paper.tex
paper

We overload \emph{substitution} as a partial operation on objects, given
a parallel substitution $\sigma$. Since the syntax of object is not
stable by (textual) substitution, it is defined \emph{hereditarily}: it
is mutually recursive with a \emph{cut} function $\gcut M S$ which can
in turn trigger a chain of substitutions, leading to a canonical objects
or an error.
\begin{align*}
  \gsubst {(\lam x M)} \sigma &= \lam x {\gsubst M \sigma} &
  \text{if $\var x\notin \dom{\sigma}$}
  \\
  \gsubst {(\app {\cst c} S)} \sigma &= \app {\cst c} {(\gsubst S
    \sigma)}
  \\
  \gsubst {(\app {\var x} S)} \sigma &=
  \app {\var x} {\gsubst S \sigma} &
  \text{if $\var x\notin \dom{\sigma}$}
  \\
  \gsubst {(\app {\var x} S)} \sigma &=
  \gcut M {\gsubst S \sigma} &
  \text{if $\sigma(\var x) = M$}
  \\
  \gsubst {\smeta X {\sigma}} {\sigma'} &=
  \smeta X {\gcomp\sigma{\sigma'}}
  \\
  \gsubst\spinenil\sigma &=
  \spinenil \\
  \gsubst{\spinecons M S}\sigma &=
  \spinecons{\gsubst M\sigma}{\gsubst S\sigma}
  \\
  \gcut {\lam x M} {\spinecons N S} &=
  \gcut {\gsubst M {\msubst x N}} S
  \\
  % TODO cas meta/S ????
  % \gcut {\smeta X\sigma} S &=
  % \smeta X{\gcomp\sigma\sigma'}
  % \\
  \gcut {\app H S} \spinenil &=
  \app H S
  \\
  \gcut {\smeta X\sigma} \spinenil &= \smeta X\sigma
  \\
  \gcomp \sigma \msubstnil &=
  \sigma
  \\
  \gcomp \sigma {(\msubstcons{\sigma'} x M)} &=
  \gcomp {\gsubst\sigma {\msubst x M}} \sigma'
  \\
  \gsubst {(\msubstcons{\sigma'} x M)} \sigma &=
  \msubstcons{\gsubst{\sigma'}\sigma} x {\gsubst M \sigma}
  \\
  \gsubst {\msubstnil}{\sigma} &= \msubstnil
\end{align*}

We extend this operation trivially on families $\gsubst A \sigma$ and
kinds $\gsubst K \sigma$. We also allow to abusively write
non-canonical applications $\app M S$ as a shorthand for $\gcut M S$
if they are defined.

@@

% TODO pourquoi cette def?
% TODO terminaison??
% TODO stabilité de la eta?
% TODO comparer avec la def usuelle

@@

  \begin{mathpar}
    \infer{ }{
      \jsig\mr\gnil\mr\gnil
    }

@@

    \infer{
      \jsig\mmr\Sigma\mmmr{\Sigma'}
      \and
      \jkind\mr\gnil {\!_{\Sigma'} K}\mmr {K'}
    }{
      \jsig\mr{\gcons\Sigma{{\cst a : K}}}\mmmr{\gcons{\Sigma'}{{\cst a : K'}}}
    }
  \end{mathpar}

@@

    \infer{
      \Sigma(a) = K
      \and
      \jfl\mr\Gamma K S \mmr {S'}
      % TODO
    }{
      \jfam\mr\Gamma {\app{\cst a} S} \mmr {\app{\cst a} {S'}}
    }
  \end{mathpar}

@@

@@

\begin{theorem}
  If \/ $\jm\mr\Gamma M A \mmr {M'}$ then \/ $\jlfm\Gamma {(\checkoutr
    M)} (\checkoutr A)$
\end{theorem}

@@

We now turn to the definition of our computation calculus. As we saw,
it is attached to special, \emph{defined} constants in the signature
and manipulates \LF\ objects. This language is un(i)typed, in that all
\LF\ objects have the same implicit type $\cst{obj}$. We have the
ability to construct objects out of other, and deconstruct them by
case analysis:
\begin{align*}
  T &\gequal
  \lam x T \gor
  U \\
  U &\gequal
  F \gor
  \matchin U \Gamma \casenobar {\app {H} {\vec x}} U\ |\ \ldots\
  \case {\app {H} {\vec x}} U
\end{align*}

@@

\section{Implementation}

\section{Related work}
% 1-2 pages

>>>>

We overload \emph{substitution} as a partial operation on objects, given
a parallel substitution $\sigma$. Since the syntax of object is not
stable by (textual) substitution, it is defined \emph{hereditarily}: it
is mutually recursive with a \emph{cut} function $\gcut M S$ which can
in turn trigger a chain of substitutions, leading to a canonical objects
or an error.
\begin{align*}
  \gsubst {(\lam x M)} \sigma &= \lam x {\gsubst M \sigma} &
  \text{if $\var x\notin \dom{\sigma}$}
  \\
  \gsubst {(\app {\cst c} S)} \sigma &= \app {\cst c} {(\gsubst S
    \sigma)}
  \\
  \gsubst {(\app {\var x} S)} \sigma &=
  \app {\var x} {\gsubst S \sigma} &
  \text{if $\var x\notin \dom{\sigma}$}
  \\
  \gsubst {(\app {\var x} S)} \sigma &=
  \gcut M {\gsubst S \sigma} &
  \text{if $\sigma(\var x) = M$}
  \\
  \gsubst {\smeta X {\sigma}} {\sigma'} &=
  \smeta X {\gcomp\sigma{\sigma'}}
  \\
  \gsubst\spinenil\sigma &=
  \spinenil \\
  \gsubst{\spinecons M S}\sigma &=
  \spinecons{\gsubst M\sigma}{\gsubst S\sigma}
  \\
  \gcut {\lam x M} {\spinecons N S} &=
  \gcut {\gsubst M {\msubst x N}} S
  \\
  % TODO cas meta/S ? rem: on en a pas besoin ds le proto.
  % \gcut {\smeta X\sigma} S &=
  % \smeta X{\gcomp\sigma\sigma'}
  % \\
  \gcut {\app H S} \spinenil &=
  \app H S
  \\
  \gcut {\smeta X\sigma} \spinenil &= \smeta X\sigma
  \\
  \gcomp \sigma \msubstnil &=
  \sigma
  \\
  \gcomp \sigma {(\msubstcons{\sigma'} x M)} &=
  \gcomp {\gsubst\sigma {\msubst x M}} \sigma'
  \\
  \gsubst {(\msubstcons{\sigma'} x M)} \sigma &=
  \msubstcons{\gsubst{\sigma'}\sigma} x {\gsubst M \sigma}
  \\
  \gsubst {\msubstnil}{\sigma} &= \msubstnil
\end{align*}

We extend this operation trivially on families $\gsubst A \sigma$ and
kinds $\gsubst K \sigma$. We also allow to abusively write
non-canonical application $\app M S$ as a shorthand for $\gcut M S$
if it is defined.

@@

% TODO pourquoi cette def?
% TODO terminaison?
% TODO stabilité de la eta?

@@

@@

    \infer{
      \jsig\mmr\Sigma\mmmr{\Sigma'}
      \and
      \jkind\mr\gnil {\!_{\Sigma'} K}\mmr {K'}
    }{
      \jsig\mr{\gcons\Sigma{{\cst a : K}}}\mmmr{\gcons{\Sigma'}{{\cst a : K'}}}
    }

    \infer{ }{
      \jsig\mr\gnil\mr\gnil
    }
  \end{mathpar}

@@

    \infer{
      \Sigma(a) = K
      \and
      \jfl\mr\Gamma K S \mmr {S'}
    }{
      \jfam\mr\Gamma {\app{\cst a} S} \mmr {\app{\cst a} {S'}}
    }
  \end{mathpar}

@@

\paragraph{Commit}
% TODO commit

@@

\begin{theorem}
  If \/ $\jm\mr\Gamma M A \mmr {M'}$ then \/ $\jlfm\Gamma {(\checkoutrr
    M)} (\checkoutrr A)$
\end{theorem}

@@

We now turn to the definition of our computation calculus. As we saw,
code is attached to special, \emph{defined} constants in the signature
that manipulates \LF\ objects. We have the ability to construct
objects out of other, and deconstruct them by case analysis:
\begin{align*}
  T &\gequal
  \lam x T \gor
  U \\
  U &\gequal
  F \gor
  \matchin U \Gamma \casenobar {\app {H} {\vec {\var x}}} U\ |\ \ldots\
  \case {\app {H} {\vec {\var x}}} U
\end{align*}

This language is un(i)typed, in that all \LF\ objects have the same
implicit type $\cst{obj}$: there is no static guarantee that terms
produced by a constant $\cst c : A_1\to A_2$ will have type $A_2$; it
is also non-terminating: we could declare a constant $\cst c : \cst
a\to \cst a = \lam x \app {\cst c} {\var x}$. We expose here only the
minimal building blocks, but other features could be added at will to
this language. You can think of defined constants as black boxes
taking and producing terms, and being able to call other black
boxes. We detail in section \ref{sec:implem} how to embed
\sysname{OCaml} as computation calculus.

Nevertheless, in order to maintain the well-typing of the generated
term, we want to maintain the invariant that arguments fed to these
defined constant as well as the computation result have the announced
type. We thus maintain \emph{dynamic typing} of computations.

@@

\section{Implementation}
\label{sec:implem}
\section{Related work}
% 1-2 pages
